---
title: "BIS15L_AI_class"
author: "Zhuoya Wang"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview
Today, we're exploring how AI can aid in writing, comprehending and troubleshooting our R code. AI is a very valuable tool, but it's important not to rely on it entirely. Building a strong R foundation is still important. These AI tools are not perfect and can be inaccurate solutions. In addition these tools can't read our minds, so understanding R helps us generate better prompts to get better responses. In short, AI should be used as a tool, not a crutch. Additionally, do not forget about other useful resources like Stack Overflow/Exchange, online tutorials, and GitHub issues pages for learning and troubleshooting. Let's combine AI with traditional methods to master R effectively.

This overview isn't exhaustive of all possible AI tools for R coding, but I believe it is a good start and covers everything you need.


**Make sure to enable Github Copilot in RStudio.**
Go to Tools -> Global Options -> Copilot -> Enable Git

## Learning Goals: 
*At the end of this exercise, you will be able to:*  
1. Explain what AI is and what AI tools can be used for R.
2. Enable Github Copilot in RStudio.
3. Use Copilot/ChatGPT to write code in RStudio.
4. Use Copilot/ChatGPT for troubleshooting code in RStudio.


## Exploration Questions:
Lets get some get a feel for what copilot (or ChatGPT) does, use Github Copilot to autocomplete the questions below. You can alternatively paste the questions in ChatGPT or Microsoft Copilot.
1. What is AI?


Artificial Intelligence (AI) refers to the branch of computer science focused on creating systems capable of performing tasks that typically require human intelligence. These tasks include understanding natural language, recognizing patterns in data, making decisions, and learning from experiences. AI technologies utilize algorithms and vast amounts of data to mimic cognitive functions such as problem-solving and learning. This field encompasses various approaches and technologies, including machine learning, deep learning, neural networks, and natural language processing, enabling applications ranging from virtual assistants and recommendation systems to autonomous vehicles and healthcare diagnostics.


2. What is ChatGPT?

ChatGPT is an advanced language model developed by OpenAI, based on the GPT (Generative Pre-trained Transformer) architecture. It is designed to generate human-like text responses by predicting subsequent words in a sequence based on the input it receives. ChatGPT has been trained on a diverse range of internet text, enabling it to understand and generate text across a wide array of topics and in various languages. It can perform tasks such as answering questions, writing essays, composing emails, generating creative stories, and more. The model's ability to process and generate natural language allows it to engage in conversations, making it a valuable tool for applications in customer service, content creation, education, and entertainment.

3. What is Github Copilot?

GitHub Copilot is an AI-powered code completion tool developed by GitHub in collaboration with OpenAI. It acts as an intelligent assistant for programmers, offering suggestions for entire lines of code or even whole functions as they type. Copilot is built on a large corpus of public source code and leverages the Codex AI model, a descendant of the GPT (Generative Pre-trained Transformer) family, to understand the context of the code being written and generate relevant, syntactically correct code snippets and documentation comments.

Designed to work within popular Integrated Development Environments (IDEs) and code editors, GitHub Copilot aims to improve developer productivity by automating routine tasks, suggesting alternative ways to solve problems, and reducing the time spent on searching for code snippets and examples. It can generate code in a variety of programming languages and for different frameworks, making it a versatile tool for software development. Despite its capabilities, developers are advised to review the suggestions it provides carefully, as it can generate incorrect or insecure code at times.



4. [Try it yourself: Insert your own question here]

## Load the libraries  
```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(janitor)
library(dplyr)
library(shiny)
library(shinydashboard)
library(shiny)
```

## Load the Data: 
```{r}
#load the Netflix Data in your Data folder, save it under the object name netflix

netflix <- read_csv("data/netflix_titles.csv")
```

*Did copilot help you write the code above? Could anything about the prompt be improved to help Copilot understand what you want to do more accurately?*


## Lets Practice:
Explore the elements in your data: include a comment within your code chunk to prompt Copilot to use the function of your choice
```{r}
glimpse(netflix)

# Summary statistics for numerical variables
# This provides a quick statistical summary of the numerical columns, including mean, median, min, max, etc.
summary(netflix)

# Count missing values in each column
# This will help identify if there are columns with many missing values that need attention
summarize_all(netflix, function(x) sum(is.na(x)))

```
*Did copilot use your preferred method to get the summary of the data? If not, how could you improve the prompt to get the desired output?*


Sort out the movies from the netflix into an new object called netflix_movies
```{r} 

```


Which director has the most movies on netflix?
```{r}

count(netflix, director, sort = T)
```
*Do you see anything wrong with the table above, look at row 1. Write a prompt below to fix this.*

Which director has the most movies on netflix? remove NA
```{r}
filter(count(netflix, director, sort = T), !is.na(director))
```



Which director had the most movies by release year and remove the NA's?
```{r}


```
*See anything else interesting about this table, look at the first director on the table. Could this be a problem and how can we edit the dataset to fix this? Try writing a prompt.*


[Insert your prompt here]
```{r}
filter(count(netflix, director, sort = T), !is.na(director))
```



Make a bar plot of the top directors with the most movies on netflix
```{r}


```
*Do you see any issues with this plot? How could we improve the prompt to get a better boxplot? Try writing a prompt.*

[Insert your prompt here]
```{r}

```

*How about for just 2020*



[Insert your prompt here]
```{r}

```


*Take a few minutes to make your own plot(s) using prompts, edit prompts until you get a plot that you like.*






## Lets make the top 10 directors barplot into a Shiny App: use copilot or Chatgpt


*Now use ChatGPT, Copilot or previous notes to fix the code*





## That's it! Let's take a break and then move on to working on group projects.
