---
title: "RMarkdown Practice"
author: "Zhuoya Wang"
date: "2024-01-11"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Title 
### Title 
#### Title
Title

[BIS 15L Webpage](https://jmledford3115.github.io/datascibiol/) ## link of web

The fat car crossed  
the road. ## two space a new line


## Practice
1. Test out the arithmetic capabilities of R; experiment by doing addition, subtraction, multiplication, and division.  

```{r}
## addition
6+8
```

```{r}
##subtraction

1004-228
```

```{r}
##multiplication

8*1230
```

```{r}
## division
1101 / 55
```


2. Go back to your "RMarkdown Practice" file and experiment with titles, text, and syntax.  

## RMarkdown Practice
RMarkdown Practice
_RMarkdown Practice_
**RMarkdown Practice**


3. Copy and paste the following two pieces of code into the document (include the gray code "chunks").      

```{r}
#install.packages("tidyverse")
library("tidyverse") 
```

```{r}
ggplot(mtcars, aes(x = factor(cyl))) +
    geom_bar()
```

4. Now `knit` the file to html using the knit button at the top of the page.  
5. Lastly, commit and push this file to your GitHub repository.  